---
title: "Hands-on Exercise 4"
author: "JIN JIAYING"
date: "Feburary 6,2026"
date-modified: "last-modified"
execute: 
  eval: true
  echo: true
  warning: false
  freeze: true
---

# 4.1 Visual Distribution

This is Hands-on Exercise 4.1 about two new statistical graphic methods for visualising distribution, ridgeline plot and raincloud plot by using ggplot2 and its extensions.

## 4.1.1 Getting Start

```{r}
pacman::p_load(ggdist, ggridges, ggthemes, colorspace, tidyverse)
```

## 4.1.2 Importing Data

```{r}
exam <- read_csv("C:/Users/jiayi/Documents/SMU/26 Jan Term/ISSS608_Visual_Analytics_and_Applications/ISSS608_Lwn0170/data/Exam_data.csv")
```

## 4.1.3 Ridgeline

```{r}
ggplot(exam, 
       aes(x = ENGLISH, 
           y = CLASS)) +
  geom_density_ridges(
    scale = 3,
    rel_min_height = 0.01,
    bandwidth = 3.4,
    fill = lighten("#7097BB", .3),
    color = "white"
  ) +
  scale_x_continuous(
    name = "English grades",
    expand = c(0, 0)
    ) +
  scale_y_discrete(name = NULL, expand = expansion(add = c(0.2, 2.6))) +
  theme_ridges()
```

### 4.1.3(1)

```{r}
ggplot(exam, 
       aes(x = ENGLISH, 
           y = CLASS,
           fill = stat(x))) +
  geom_density_ridges_gradient(
    scale = 3,
    rel_min_height = 0.01) +
  scale_fill_viridis_c(name = "Temp. [F]",
                       option = "C") +
  scale_x_continuous(
    name = "English grades",
    expand = c(0, 0)
  ) +
  scale_y_discrete(name = NULL, expand = expansion(add = c(0.2, 2.6))) +
  theme_ridges()
```

### 4.1.3(2)

```{r}
ggplot(exam,
       aes(x = ENGLISH, 
           y = CLASS, 
           fill = 0.5 - abs(0.5-stat(ecdf)))) +
  stat_density_ridges(geom = "density_ridges_gradient", 
                      calc_ecdf = TRUE) +
  scale_fill_viridis_c(name = "Tail probability",
                       direction = -1) +
  theme_ridges()
```

### 4.1.3(3)

```{r}
ggplot(exam,
       aes(x = ENGLISH, 
           y = CLASS, 
           fill = factor(stat(quantile))
           )) +
  stat_density_ridges(
    geom = "density_ridges_gradient",
    calc_ecdf = TRUE, 
    quantiles = 4,
    quantile_lines = TRUE) +
  scale_fill_viridis_d(name = "Quartiles") +
  theme_ridges()
```

### 4.1.3(4)

```{r}
ggplot(exam,
       aes(x = ENGLISH, 
           y = CLASS, 
           fill = factor(stat(quantile))
           )) +
  stat_density_ridges(
    geom = "density_ridges_gradient",
    calc_ecdf = TRUE, 
    quantiles = c(0.025, 0.975)
    ) +
  scale_fill_manual(
    name = "Probability",
    values = c("#FF0000A0", "#A0A0A0A0", "#0000FFA0"),
    labels = c("(0, 0.025]", "(0.025, 0.975]", "(0.975, 1]")
  ) +
  theme_ridges()
```

## 4.1.4 Raincloud Plot

```{r}
ggplot(exam, 
       aes(x = RACE, 
           y = ENGLISH)) +
  stat_halfeye(adjust = 0.5,
               justification = -0.2,
               .width = 0,
               point_colour = NA)
```

### 4.1.4(1)

```{r}
ggplot(exam, 
       aes(x = RACE, 
           y = ENGLISH)) +
  stat_halfeye(adjust = 0.5,
               justification = -0.2,
               .width = 0,
               point_colour = NA) +
  geom_boxplot(width = .20,
               outlier.shape = NA)
```

### 4.1.4(2)

```{r}
ggplot(exam, 
       aes(x = RACE, 
           y = ENGLISH)) +
  stat_halfeye(adjust = 0.5,
               justification = -0.2,
               .width = 0,
               point_colour = NA) +
  geom_boxplot(width = .20,
               outlier.shape = NA) +
  stat_dots(side = "left", 
            justification = 1.2, 
            binwidth = .5,
            dotsize = 2)
```

### 4.1.4(3)

```{r}
ggplot(exam, 
       aes(x = RACE, 
           y = ENGLISH)) +
  stat_halfeye(adjust = 0.5,
               justification = -0.2,
               .width = 0,
               point_colour = NA) +
  geom_boxplot(width = .20,
               outlier.shape = NA) +
  stat_dots(side = "left", 
            justification = 1.2, 
            binwidth = .5,
            dotsize = 1.5) +
  coord_flip() +
  theme_economist()
```

# 4.2 Visual Statistical Analysis

This is Hands-on Exercise 4.2 about ggstatsplot, performance and parameter packages to create visual graphics and visualise model.

## 4.2.1 Getting Start

```{r}
pacman::p_load(ggstatsplot, tidyverse)
```

## 4.2.2 Importing Data

Same as 4.1.2

## 4.2.3 Analysis Ploting

### 4.2.3(1)

```{r}
set.seed(1234)

gghistostats(
  data = exam,
  x = ENGLISH,
  type = "bayes",
  test.value = 60,
  xlab = "English scores"
)
```

### 4.2.3(2)

```{r}
ggbetweenstats(
  data = exam,
  x = GENDER, 
  y = MATHS,
  type = "np",
  messages = FALSE
)
```

### 4.2.3(3)

```{r}
ggbetweenstats(
  data = exam,
  x = RACE, 
  y = ENGLISH,
  type = "p",
  mean.ci = TRUE, 
  pairwise.comparisons = TRUE, 
  pairwise.display = "s",
  p.adjust.method = "fdr",
  messages = FALSE
)
```

### 4.2.3(4)

```{r}
ggscatterstats(
  data = exam,
  x = MATHS,
  y = ENGLISH,
  marginal = FALSE,
  )
```

### 4.2.3(5)

```{r}
exam1 <- exam %>% 
  mutate(MATHS_bins = 
           cut(MATHS, 
               breaks = c(0,60,75,85,100))
)
ggbarstats(exam1, 
           x = MATHS_bins, 
           y = GENDER)
```

# 4.3 Visualising Uncertainty

This is Hands-on Exercise 4.3 about creating statistical graphics for visualising uncertainty.

## 4.3.1 Getting Start

```{r}
pacman::p_load(scales, viridis, lubridate, ggthemes,
               gridExtra, readxl, knitr, data.table,
               CGPfunctions, ggHoriPlot, tidyverse, crosstalk,
               plotly, ggforce)
```

## 4.3.2 Importing Data

Same as 4.1.2 \## 4.3.3 ggplot2 method

```{r}
my_sum <- exam %>%
  group_by(RACE) %>%
  summarise(
    n=n(),
    mean=mean(MATHS),
    sd=sd(MATHS)
    ) %>%
  mutate(se=sd/sqrt(n-1))
knitr::kable(head(my_sum), format = 'html')
```

### 4.3.3(1)

```{r}
ggplot(my_sum) +
  geom_errorbar(
    aes(x=RACE, 
        ymin=mean-se, 
        ymax=mean+se), 
    width=0.2, 
    colour="black", 
    alpha=0.9, 
    linewidth=0.5) +
  geom_point(aes
           (x=RACE, 
            y=mean), 
           stat="identity", 
           color="red",
           size = 1.5,
           alpha=1) +
  ggtitle("Standard error of mean maths score by rac")
```

### 4.3.3(2)

```{r}
ggplot(my_sum) +
  geom_errorbar(
    aes(x=reorder(RACE, -mean), 
        ymin=mean-1.96*se, 
        ymax=mean+1.96*se), 
    width=0.2, 
    colour="black", 
    alpha=0.9, 
    linewidth=0.5) +
  geom_point(aes
           (x=RACE, 
            y=mean), 
           stat="identity", 
           color="red",
           size = 1.5,
           alpha=1) +
  labs(x = "Maths score",
       title = "95% confidence interval of mean maths score by race")
```

### 4.3.3(3)

```{r}
my_sum2 <- my_sum %>%
  mutate(across(c(mean, sd, se), ~ round(.x, 2)))

shared_df <- SharedData$new(my_sum2, key = ~RACE)


bscols(
  widths = c(4, 8),

  ggplotly(
    ggplot(shared_df) +
      geom_errorbar(
        aes(
          x = reorder(RACE, -mean),
          ymin = mean - 2.58 * se,
          ymax = mean + 2.58 * se
        ),
        width  = 0.2,
        colour = "black",
        alpha  = 0.9
      ) +
      geom_point(
        aes(
          x = RACE,
          y = mean,
          text = paste(
            "Race:", RACE,
            "<br>N:", n,
            "<br>Avg. Scores:", mean,   # 这里已经 round 过了
            "<br>99% CI:[",
            round(mean - 2.58 * se, 2), ",",
            round(mean + 2.58 * se, 2), "]"
          )
        ),
        color = "red",
        size  = 1.5,
        alpha = 1
      ) +
      xlab("Race") +
      ylab("Average Scores") +
      theme_minimal() +
      theme(axis.text.x = element_text(angle = 45, vjust = 0.5, hjust = 1)) +
      ggtitle("99% Confidence interval of average /<br>maths scores by race"),
    tooltip = "text"
  ),

  DT::datatable(
    shared_df,
    rownames = FALSE,
    class = "compact",
    width = "100%",
    options = list(pageLength = 10, scrollX = TRUE),
    colnames = c("Race", "No. of pupils", "Avg Scores", "Std Dev", "Std Error")
  )
)

```

## 4.3.4 ggdist package

```{r}
exam %>%
  ggplot(aes(x = RACE, 
             y = MATHS)) +
  stat_pointinterval() +
  labs(
    title = "Visualising confidence intervals of mean math score",
    subtitle = "Mean Point + Multiple-interval plot")

exam %>%
  ggplot(aes(x = RACE, y = MATHS)) +
  stat_pointinterval(.width = 0.95,
  .point = median,
  .interval = qi) +
  labs(
    title = "Visualising confidence intervals of median math score",
    subtitle = "Median Point + Multiple-interval plot")
```

### 4.3.4(1)

```{r}
exam %>%
  ggplot(aes(x = RACE, 
             y = MATHS)) +
  stat_pointinterval(
    show.legend = FALSE) +   
  labs(
    title = "Visualising confidence intervals of mean math score",
    subtitle = "Mean Point + Multiple-interval plot")
```

### 4.3.4(2)

```{r}
exam %>%
  ggplot(aes(x = RACE, 
             y = MATHS)) +
  stat_gradientinterval(   
    fill = "skyblue",      
    show.legend = TRUE     
  ) +                        
  labs(
    title = "Visualising confidence intervals of mean math score",
    subtitle = "Gradient + interval plot")
```

## 4.3.5 Hypothetical Outcome Plots (HOPs)

```{r}
library(ggplot2)
library(ggforce)
library(gganimate)
library(ungeviz)
```

```{r}
ggplot(data = exam, 
       (aes(x = factor(RACE), 
            y = MATHS))) +
  geom_point(position = position_jitter(
    height = 0.3, 
    width = 0.05), 
    size = 0.4, 
    color = "#0072B2", 
    alpha = 1/2) +
  geom_hpline(data = sampler(25, 
                             group = RACE), 
              height = 0.6, 
              color = "#D55E00") +
  theme_bw() + 
  transition_states(.draw, 1, 3)
```

# 4.4 Funnel Plots for Fair Comparisons

This is Hands-on Exercise 4.4 about plotting static and interactive funnel plots.

## 4.4.1 Getting Start

```{r}
pacman::p_load(tidyverse, FunnelPlotR, plotly, knitr)
```

## 4.4.2 Importing Data

```{r}
covid19 <- read_csv("C:/Users/jiayi/Documents/SMU/26 Jan Term/ISSS608_Visual_Analytics_and_Applications/ISSS608_Lwn0170/data/COVID-19_DKI_Jakarta.csv")%>%
  mutate_if(is.character, as.factor)
```

## 4.4.3 FunnelPlotR methods

### 4.4.3(1)

```{r}
funnel_plot(
  .data = covid19,
  numerator = Positive,
  denominator = Death,
  group = `Sub-district`
)
```

### 4.4.3(2)

```{r}
funnel_plot(
  .data = covid19,
  numerator = Death,
  denominator = Positive,
  group = `Sub-district`,
  data_type = "PR",     #<<
  xrange = c(0, 6500),  #<<
  yrange = c(0, 0.05)   #<<
)
```

### 4.4.3(3)

```{r}
funnel_plot(
  .data = covid19,
  numerator = Death,
  denominator = Positive,
  group = `Sub-district`,
  data_type = "PR",   
  xrange = c(0, 6500),  
  yrange = c(0, 0.05),
  label = NA,
  title = "Cumulative COVID-19 Fatality Rate by Cumulative Total Number of COVID-19 Positive Cases", #<<           
  x_label = "Cumulative COVID-19 Positive Cases", #<<
  y_label = "Cumulative Fatality Rate"  #<<
)
```

## 4.4.4 gplot2 methods

```{r}
df <- covid19 %>%
  mutate(rate = Death / Positive) %>%
  mutate(rate.se = sqrt((rate*(1-rate)) / (Positive))) %>%
  filter(rate > 0)
fit.mean <- weighted.mean(df$rate, 1/df$rate.se^2)

number.seq <- seq(1, max(df$Positive), 1)
number.ll95 <- fit.mean - 1.96 * sqrt((fit.mean*(1-fit.mean)) / (number.seq)) 
number.ul95 <- fit.mean + 1.96 * sqrt((fit.mean*(1-fit.mean)) / (number.seq)) 
number.ll999 <- fit.mean - 3.29 * sqrt((fit.mean*(1-fit.mean)) / (number.seq)) 
number.ul999 <- fit.mean + 3.29 * sqrt((fit.mean*(1-fit.mean)) / (number.seq)) 
dfCI <- data.frame(number.ll95, number.ul95, number.ll999, 
                   number.ul999, number.seq, fit.mean)
```

### 4.4.4(1)

```{r}
p <- ggplot(df, aes(x = Positive, y = rate)) +
  geom_point(aes(label=`Sub-district`), 
             alpha=0.4) +
  geom_line(data = dfCI, 
            aes(x = number.seq, 
                y = number.ll95), 
            size = 0.4, 
            colour = "grey40", 
            linetype = "dashed") +
  geom_line(data = dfCI, 
            aes(x = number.seq, 
                y = number.ul95), 
            size = 0.4, 
            colour = "grey40", 
            linetype = "dashed") +
  geom_line(data = dfCI, 
            aes(x = number.seq, 
                y = number.ll999), 
            size = 0.4, 
            colour = "grey40") +
  geom_line(data = dfCI, 
            aes(x = number.seq, 
                y = number.ul999), 
            size = 0.4, 
            colour = "grey40") +
  geom_hline(data = dfCI, 
             aes(yintercept = fit.mean), 
             size = 0.4, 
             colour = "grey40") +
  coord_cartesian(ylim=c(0,0.05)) +
  annotate("text", x = 1, y = -0.13, label = "95%", size = 3, colour = "grey40") + 
  annotate("text", x = 4.5, y = -0.18, label = "99%", size = 3, colour = "grey40") + 
  ggtitle("Cumulative Fatality Rate by Cumulative Number of COVID-19 Cases") +
  xlab("Cumulative Number of COVID-19 Cases") + 
  ylab("Cumulative Fatality Rate") +
  theme_light() +
  theme(plot.title = element_text(size=12),
        legend.position = c(0.91,0.85), 
        legend.title = element_text(size=7),
        legend.text = element_text(size=7),
        legend.background = element_rect(colour = "grey60", linetype = "dotted"),
        legend.key.height = unit(0.3, "cm"))
p
```

### 4.4.4(2)

```{r}
fp_ggplotly <- ggplotly(p,
  tooltip = c("label", 
              "x", 
              "y"))
fp_ggplotly
```
