---
title: "Hands-on Exercise 10"
author: "JIN JIAYING"
date: "March 25,2026"
date-modified: "last-modified"
execute: 
  eval: true
  echo: true
  warning: false
  freeze: true
---

# 10 Information Dashboard Design: R methods

This is Hands-on Exercise 10 to 

- create bullet chart by using ggplot2
- create sparklines by using ggplot2
- build industry standard dashboard by using R Shiny

# 10.1 Getting Start
```{r}
pacman::p_load(lubridate, ggthemes, reactable, reactablefmtr, gt, gtExtras, tidyverse, RODBC, dataui, reactablefmtr)
```

# 10.2 Data Importing
```{r}
con <- odbcConnectAccess2007('data/Coffee Chain.mdb')
coffeechain <- sqlFetch(con, 'CoffeeChain Query')
write_rds(coffeechain, "data/CoffeeChain.rds")
odbcClose(con)
```

# 10.3 Data Preparation
```{r}
coffeechain <- read_rds("data/CoffeeChain.rds")

product <- coffeechain %>%
  group_by(`Product`) %>%
  summarise(`target` = sum(`Budget Sales`), `current` = sum(`Sales`)) %>%
  ungroup()
```

# 10.4 Bullet chart in ggplot2
```{r}
ggplot(product, aes(Product, current)) + 
  geom_col(aes(Product, max(target) * 1.01), fill="grey85", width=0.85) +
  geom_col(aes(Product, target * 0.75), fill="grey60", width=0.85) +
  geom_col(aes(Product, target * 0.5), fill="grey50", width=0.85) +
  geom_col(aes(Product, current), width=0.35, fill = "black") + 
  geom_errorbar(aes(y = target, x = Product, ymin = target, ymax= target), width = .4, colour = "red", size = 1) +
  coord_flip()
```

# 10.5 Plotting sparklines using ggplot2
```{r}
sales_report <- coffeechain %>%
  filter(Date >= "2013-01-01") %>%
  mutate(Month = month(Date)) %>%
  group_by(Month, Product) %>%
  summarise(Sales = sum(Sales)) %>%
  ungroup() %>%
  select(Month, Product, Sales)

mins <- group_by(sales_report, Product) %>%
  slice(which.min(Sales))
maxs <- group_by(sales_report, Product) %>%
  slice(which.max(Sales))
ends <- group_by(sales_report, Product) %>% 
  filter(Month == max(Month))

quarts <- sales_report %>%
  group_by(Product) %>%
  summarise(quart1 = quantile(Sales, 0.25),
            quart2 = quantile(Sales, 0.75)) %>%
  right_join(sales_report)

ggplot(sales_report, aes(x=Month, y=Sales)) + 
  facet_grid(Product ~ ., scales = "free_y") + 
  geom_ribbon(data = quarts, aes(ymin = quart1, max = quart2), fill = 'grey90') +
  geom_line(size=0.3) +
  geom_point(data = mins, col = 'red') +
  geom_point(data = maxs, col = 'blue') +
  geom_text(data = mins, aes(label = Sales), vjust = -1) +
  geom_text(data = maxs, aes(label = Sales), vjust = 2.5) +
  geom_text(data = ends, aes(label = Sales), hjust = 0, nudge_x = 0.5) +
  geom_text(data = ends, aes(label = Product), hjust = 0, nudge_x = 1.0) +
  expand_limits(x = max(sales_report$Month) + (0.25 * (max(sales_report$Month) - min(sales_report$Month)))) +
  scale_x_continuous(breaks = seq(1, 12, 1)) +
  scale_y_continuous(expand = c(0.1, 0)) +
  theme_tufte(base_size = 3, base_family = "Helvetica") +
  theme(axis.title=element_blank(), axis.text.y = element_blank(), 
        axis.ticks = element_blank(), strip.text = element_blank())
```

# 10.6 Static Information Dashboard Design: gt and gtExtras methods
```{r}
product %>%
  gt::gt() %>%
  gt_plt_bullet(column = current, target = target, width = 60, palette = c("lightblue", "black")) %>%
  gt_theme_538()
```

# 10.7 sparklines: gtExtras method
```{r}
report <- coffeechain %>%
  mutate(Year = year(Date)) %>%
  filter(Year == "2013") %>%
  mutate (Month = month(Date, label = TRUE, abbr = TRUE)) %>%
  group_by(Product, Month) %>%
  summarise(Sales = sum(Sales)) %>%
  ungroup()

report %>%
  group_by(Product) %>%
  summarize('Monthly Sales' = list(Sales), 
            .groups = "drop")
```
```{r}
report %>%
  group_by(Product) %>%
  summarize('Monthly Sales' = list(Sales), 
            .groups = "drop") %>%
   gt() %>%
   gt_plt_sparkline('Monthly Sales', same_limit = FALSE)
```
```{r}
report %>% 
  group_by(Product) %>% 
  summarise("Min" = min(Sales, na.rm = T),
            "Max" = max(Sales, na.rm = T),
            "Average" = mean(Sales, na.rm = T)) %>%
  gt() %>%
  fmt_number(columns = 4, decimals = 2)
```
```{r}
spark <- report %>%
  group_by(Product) %>%
  summarize('Monthly Sales' = list(Sales), 
            .groups = "drop")

sales <- report %>% 
  group_by(Product) %>% 
  summarise("Min" = min(Sales, na.rm = T),
            "Max" = max(Sales, na.rm = T),
            "Average" = mean(Sales, na.rm = T))

sales_data = left_join(sales, spark)
```
```{r}
sales_data %>%
  gt() %>%
  gt_plt_sparkline('Monthly Sales', same_limit = FALSE)
```
```{r}
bullet <- coffeechain %>%
  filter(Date >= "2013-01-01") %>%
  group_by(`Product`) %>%
  summarise(`Target` = sum(`Budget Sales`), `Actual` = sum(`Sales`)) %>%
  ungroup() 

sales_data = sales_data %>%
  left_join(bullet)

sales_data %>%
  gt() %>%
  gt_plt_sparkline('Monthly Sales') %>%
  gt_plt_bullet(column = Actual, target = Target, width = 28, palette = c("lightblue", "black")) %>%
  gt_theme_538()
```

# 10.8 Interactive Information Dashboard Design: reactable and reactablefmtr methods
```{r}
report <- report %>%
  group_by(Product) %>%
  summarize(`Monthly Sales` = list(Sales))

reactable(report, columns = list(Product = colDef(maxWidth = 200), `Monthly Sales` = colDef(cell = react_sparkline(report))))
```
```{r}
reactable(report, defaultPageSize = 13, columns = list(Product = colDef(maxWidth = 200), `Monthly Sales` = colDef(cell = react_sparkline(report))))
```
```{r}
reactable(report, defaultPageSize = 13, columns = list(Product = colDef(maxWidth = 200), `Monthly Sales` = colDef(cell = react_sparkline(report, highlight_points = highlight_points(min = "red", max = "blue"), labels = c("first", "last")))))
```
```{r}
reactable(report, defaultPageSize = 13, columns = list(Product = colDef(maxWidth = 200), `Monthly Sales` = colDef(cell = react_sparkline(report, highlight_points = highlight_points(min = "red", max = "blue"), statline = "mean"))))
```
```{r}
reactable(report, defaultPageSize = 13, columns = list(Product = colDef(maxWidth = 200), `Monthly Sales` = colDef(cell = react_sparkline(report, highlight_points = highlight_points(min = "red", max = "blue"), line_width = 1, bandline = "innerquartiles", bandline_color = "green"))))
```
```{r}
reactable(report, defaultPageSize = 13, columns = list(Product = colDef(maxWidth = 200), `Monthly Sales` = colDef(cell = react_sparkbar(report, highlight_bars = highlight_bars(min = "red", max = "blue"), bandline = "innerquartiles", statline = "mean"))))
```
