---
title: "Take-home Exercise 2"
author: "JIN JIAYING"
date: "March 6,2026"
date-modified: "last-modified"
execute: 
  eval: true
  echo: true
  warning: false
  freeze: true
---

# Overview

This is Take-home Exercise 2 based on 2 complementary datasets from the COFINFAD (Colombian Fintech Financial Analytics Dataset) context: a customer-level dataset containing demographic, product adoption, engagement, and satisfaction-related attributes, and a transaction-level dataset capturing detailed customer financial activities. Together, these datasets provide both a high-level view of customer characteristics and a more granular view of transaction behavior, making them well suited for interactive visual analysis.

The proposed Shiny prototype focuses on Exploratory Data Analysis (EDA) and Comparative Data Analysis (CDA) as its core analytical modules. EDA is used to understand the overall structure, quality, and distribution of variables across both datasets, while CDA is used to compare meaningful customer groups based on characteristics such as churn risk, customer lifetime value, engagement level, and product adoption. The inclusion of transaction-level data further strengthens the analysis by enabling comparisons of transaction frequency, value, timing, and behavioral patterns across customer segments.

To support these analytical goals, the project requires a structured data preparation process. This includes cleaning the customer and transaction datasets separately, standardizing variable types, addressing missingness, validating linking keys, and aggregating transaction-level records into customer-level behavioral summaries. The resulting analysis-ready dataset provides the foundation for a Shiny prototype that delivers interpretable, business-relevant, and visually interactive insights into customer behavior and financial activity.

# Getting Preparation
```{r}
pacman::p_load(tidyverse, skimr)
```

# Importing Data
```{r}
customer_raw <- read_csv("Data/TE02/customer_data.csv")

transaction_raw <- read_csv("Data/TE02/transactions_data.csv")
```

# Data Preparation

Behavioral and transactional data from 48,723 customers of a Colombian fintech company, collected over 12 months from January 4, 2023, to December 29, 2023. Comprises 3,159,157 individual transactions

Task:
- check identity of customer_id and transaction_id
- check if can use customer_id as join key (fully match)
- clean customer data
  -- type conversion
  -- derived variables (factor standardization, ordered factor)
  -- missing value
  -- create churn_group, clv_group, tenure_bucket etc.
- clean transaction data
  -- datetime convert to transaction_month, transaction_weekday, transaction_hour, is_weekend
  -- derived variable (factor standardization, ordered factor)
  -- transaction_amount (numeric? negative? 0 value? extreme? currency?)
  -- check duplicated
  -- missing value
- customer-level aggregated features
- create analysis_data

## Customer Data Preparation
```{r}
# Check Missing and Uniquness
customer_raw %>%
  summarise(across(everything(), ~ sum(is.na(.)))) %>%
  pivot_longer(cols = everything(),
               names_to = "variable",
               values_to = "missing_count") %>%
  arrange(desc(missing_count))

customer_raw %>%
  summarise(
    total_rows = n(),
    unique_customers = n_distinct(customer_id),
    duplicated_customer_id = n() - n_distinct(customer_id)
  )
```
```{r}
#| label: clean-customer-data

customer_clean <- customer_raw %>%
  mutate(
    # Convert date columns
    first_tx = ymd(first_tx),
    last_tx = ymd(last_tx),
    last_survey_date = ymd(last_survey_date),
    last_transaction_date = ymd(last_transaction_date),
    first_transaction_date = ymd(first_transaction_date),

    # Convert categorical variables
    gender = as.factor(gender),
    location = as.factor(location),
    income_bracket = factor(income_bracket, ordered = TRUE),
    occupation = as.factor(occupation),
    education_level = factor(education_level, ordered = TRUE),
    marital_status = as.factor(marital_status),
    acquisition_channel = as.factor(acquisition_channel),
    customer_segment = as.factor(customer_segment),
    clv_segment = factor(clv_segment, ordered = TRUE),
    feedback_sentiment = as.factor(feedback_sentiment),
    preferred_transaction_type = as.factor(preferred_transaction_type),

    # Convert logical / yes-no style fields
    savings_account = as.factor(savings_account),
    credit_card = as.factor(credit_card),
    personal_loan = as.factor(personal_loan),
    investment_account = as.factor(investment_account),
    insurance_product = as.factor(insurance_product),
    bill_payment_user = as.factor(bill_payment_user),
    auto_savings_enabled = as.factor(auto_savings_enabled),

    # Handle missing categorical values
    complaint_topics = fct_explicit_na(as.factor(complaint_topics),
                                       na_level = "No Complaint / Missing"),
    feature_requests = fct_explicit_na(as.factor(feature_requests),
                                       na_level = "No Request / Missing")
  ) %>%
  mutate(
    # Derived variables
    churn_group = case_when(
      churn_probability < 0.20 ~ "Low",
      churn_probability < 0.40 ~ "Medium",
      TRUE ~ "High"
    ),
    churn_group = factor(churn_group,
                         levels = c("Low", "Medium", "High"),
                         ordered = TRUE),

    clv_group = ntile(customer_lifetime_value, 3) %>%
      factor(levels = 1:3,
             labels = c("Low CLV", "Medium CLV", "High CLV"),
             ordered = TRUE),

    activity_group = ntile(app_logins_frequency, 3) %>%
      factor(levels = 1:3,
             labels = c("Low Activity", "Medium Activity", "High Activity"),
             ordered = TRUE),

    tenure_bucket = case_when(
      customer_tenure < 6 ~ "New",
      customer_tenure < 18 ~ "Growing",
      TRUE ~ "Mature"
    ),
    tenure_bucket = factor(tenure_bucket,
                           levels = c("New", "Growing", "Mature"),
                           ordered = TRUE),

    complaint_flag = if_else(complaint_topics == "No Complaint / Missing", "No", "Yes"),
    request_flag = if_else(feature_requests == "No Request / Missing", "No", "Yes")
  )
```
```{r}
customer_clean %>%
  select(where(is.numeric)) %>%
  skim()
```

## Transaction Data Preparation
```{r}
# Check Missing
transaction_raw %>%
  summarise(across(everything(), ~ sum(is.na(.)))) %>%
  pivot_longer(cols = everything(),
               names_to = "variable",
               values_to = "missing_count") %>%
  arrange(desc(missing_count))
```

# Exploratory Data Analysis (EDA)

## Objective of EDA

The purpose of Exploratory Data Analysis (EDA) in this exercise is to understand the overall structure, quality, and distribution of the available data before building the Shiny prototype. EDA helps identify the main characteristics of customers and transactions, detect unusual patterns or potential data issues, and highlight variables that are most suitable for interactive visualization.

## EDA on Customer-Level Data

Customer-level EDA focuses on understanding the distribution of customer attributes such as demographics, product adoption, engagement, satisfaction, and business outcomes. This includes examining variables such as age, gender, location, income bracket, customer segment, number of active products, app usage, satisfaction score, NPS, churn probability, and customer lifetime value. The goal is to reveal the overall composition of the customer base and identify key patterns worth further comparison.

```{r}

```

## EDA on Transaction-Level Data

Transaction-level EDA focuses on the behavioral side of the data. This includes exploring transaction frequency, transaction amount distributions, transaction categories, time-based patterns, and transaction status. Depending on the available fields, the analysis may also examine daily, weekly, or monthly transaction trends, differences between weekday and weekend behavior, and the prevalence of failed or international transactions. This helps provide a more detailed understanding of how customers interact with fintech services.

```{r}

```


# Comparative Data Analysis (CDA)

## Objective of CDA

The purpose of Comparative Data Analysis (CDA) is to move beyond general exploration and examine how customer groups differ across important behavioral and business-related dimensions. While EDA provides an overall view of the data, CDA focuses on structured comparisons that can generate more targeted insights for customer analysis and dashboard interaction.

## Customer Group Comparisons

CDA can be used to compare customer groups based on characteristics such as churn risk, customer lifetime value segment, product adoption, engagement level, or demographic attributes. For example, the analysis may compare whether customers with higher churn probability show lower activity, lower satisfaction, fewer active products, or different transaction patterns than lower-risk customers. Similar comparisons can also be made across income groups, customer segments, or geographic locations.

```{r}

```

## Transaction Behavior Comparisons

With the inclusion of transaction-level data, CDA can also compare transaction behavior across different customer groups. This may include differences in transaction frequency, average transaction value, failed transaction ratio, use of transaction types, or weekend transaction behavior. These comparisons help connect customer profiles with actual financial behavior, making the analysis more informative and business-relevant.

```{r}

```


# Shiny Prototype



