---
title: "Take-home Exercise 2 - 1"
author: "JIN JIAYING"
date: "March 15,2026"
date-modified: "last-modified"
execute: 
  eval: true
  echo: true
  warning: false
  freeze: true
---

# Overview

This is another part of Take-home Exercise 2 including EDA and CDA which will not be used for group project prototype building.

# Getting Preparation
```{r}
pacman::p_load(tidyverse, skimr, scales, corrplot, DT, janitor, forcats, patchwork)
```

# Importing Data
```{r}
customer_raw <- read_csv("Data/TE02/customer_data.csv")

transaction_raw <- read_csv("Data/TE02/transactions_data.csv")
```

# Data Preparation

Behavioral and transactional data from 48,723 customers of a Colombian fintech company, collected over 12 months from January 4, 2023, to December 29, 2023. Comprises 3,159,157 individual transactions

## Customer Data Preparation
```{r}
# Check Missing and Uniquness
customer_raw %>%
  summarise(across(everything(), ~ sum(is.na(.)))) %>%
  pivot_longer(cols = everything(),
               names_to = "variable",
               values_to = "missing_count") %>%
  arrange(desc(missing_count))

customer_raw %>%
  summarise(
    total_rows = n(),
    unique_customers = n_distinct(customer_id),
    duplicated_customer_id = n() - n_distinct(customer_id)
  )
```
```{r}
# clean-customer-data

customer_clean <- customer_raw %>%
  mutate(
    # Convert date columns
    first_tx = ymd(first_tx),
    last_tx = ymd(last_tx),
    last_survey_date = ymd(last_survey_date),
    last_transaction_date = ymd(last_transaction_date),
    first_transaction_date = ymd(first_transaction_date),

    # Convert categorical variables
    gender = as.factor(gender),
    location = as.factor(location),
    income_bracket = factor(income_bracket, ordered = TRUE),
    occupation = as.factor(occupation),
    education_level = factor(education_level, ordered = TRUE),
    marital_status = as.factor(marital_status),
    acquisition_channel = as.factor(acquisition_channel),
    customer_segment = as.factor(customer_segment),
    clv_segment = factor(clv_segment, ordered = TRUE),
    feedback_sentiment = as.factor(feedback_sentiment),
    preferred_transaction_type = as.factor(preferred_transaction_type),

    # Convert product ownership fields to numeric 0/1
    savings_account = case_when(
      as.character(savings_account) == "TRUE"  ~ 1,
      as.character(savings_account) == "FALSE" ~ 0,
      TRUE ~ NA_real_
    ),
    credit_card = case_when(
      as.character(credit_card) == "TRUE"  ~ 1,
      as.character(credit_card) == "FALSE" ~ 0,
      TRUE ~ NA_real_
    ),
    personal_loan = case_when(
      as.character(personal_loan) == "TRUE"  ~ 1,
      as.character(personal_loan) == "FALSE" ~ 0,
      TRUE ~ NA_real_
    ),
    investment_account = case_when(
      as.character(investment_account) == "TRUE"  ~ 1,
      as.character(investment_account) == "FALSE" ~ 0,
      TRUE ~ NA_real_
    ),
    insurance_product = case_when(
      as.character(insurance_product) == "TRUE"  ~ 1,
      as.character(insurance_product) == "FALSE" ~ 0,
      TRUE ~ NA_real_
    ),
    bill_payment_user = case_when(
      as.character(bill_payment_user) == "TRUE"  ~ 1,
      as.character(bill_payment_user) == "FALSE" ~ 0,
      TRUE ~ NA_real_
    ),
    auto_savings_enabled = case_when(
      as.character(auto_savings_enabled) == "TRUE"  ~ 1,
      as.character(auto_savings_enabled) == "FALSE" ~ 0,
      TRUE ~ NA_real_
    ),

    # Handle missing categorical values
    complaint_topics = fct_explicit_na(as.factor(complaint_topics),
                                       na_level = "No Complaint / Missing"),
    feature_requests = fct_explicit_na(as.factor(feature_requests),
                                       na_level = "No Request / Missing")
  ) %>%
  mutate(
    # Derived variables
    churn_group = case_when(
      churn_probability < 0.20 ~ "Low",
      churn_probability < 0.40 ~ "Medium",
      TRUE ~ "High"
    ),
    churn_group = factor(churn_group,
                         levels = c("Low", "Medium", "High"),
                         ordered = TRUE),

    clv_group = ntile(customer_lifetime_value, 3) %>%
      factor(levels = 1:3,
             labels = c("Low CLV", "Medium CLV", "High CLV"),
             ordered = TRUE),

    activity_group = ntile(app_logins_frequency, 3) %>%
      factor(levels = 1:3,
             labels = c("Low Activity", "Medium Activity", "High Activity"),
             ordered = TRUE),

    tenure_bucket = case_when(
      customer_tenure < 6 ~ "New",
      customer_tenure < 18 ~ "Growing",
      TRUE ~ "Mature"
    ),
    tenure_bucket = factor(tenure_bucket,
                           levels = c("New", "Growing", "Mature"),
                           ordered = TRUE),

    complaint_flag = if_else(complaint_topics == "No Complaint / Missing", "No", "Yes"),
    request_flag = if_else(feature_requests == "No Request / Missing", "No", "Yes"),
    
    # Add age group for later EDA/CDA
    age_group = case_when(
      age < 25 ~ "<25",
      age >= 25 & age <= 34 ~ "25-34",
      age >= 35 & age <= 44 ~ "35-44",
      age >= 45 & age <= 54 ~ "45-54",
      age >= 55 ~ "55+",
      TRUE ~ NA_character_
    ),
    age_group = factor(
      age_group,
      levels = c("<25", "25-34", "35-44", "45-54", "55+"),
      ordered = TRUE
    ),
    
    # Count number of products owned
    total_products_5 = rowSums(
      pick(
        savings_account,
        credit_card,
        personal_loan,
        investment_account,
        insurance_product
      ),
      na.rm = TRUE
    )
  )
```
## Transaction Data Preparation
```{r}
# Check Missing
transaction_raw %>%
  summarise(across(everything(), ~ sum(is.na(.)))) %>%
  pivot_longer(cols = everything(),
               names_to = "variable",
               values_to = "missing_count") %>%
  arrange(desc(missing_count))
```
```{r}
# Check whether each customer_id in transaction_csv can be found in customer_csv
all(transaction_raw$customer_id %in% customer_clean$customer_id)
```
The result states every customer_id in transaction data can be found in customer data.

```{r}
# clean-transaction-data
transaction_clean <- transaction_raw %>%
  mutate(
    transaction_date = ymd(date),
    transaction_day = as_date(transaction_date),
    transaction_month = floor_date(transaction_day, unit = "month"),
    transaction_weekday = wday(transaction_day, label = TRUE, abbr = FALSE),
    transaction_hour = hour(transaction_date),
    is_weekend = if_else(transaction_weekday %in%
    c("Saturday", "Sunday"), "Weekend", "Weekday"),
    transaction_type = as.factor(type)
  ) %>%
  distinct()

# key matching
transaction_clean %>%
  anti_join(customer_clean, by = "customer_id") %>%
  summarise(unmatched_customer_ids = n_distinct(customer_id))
```
```{r}
# aggregate transaction into customer level
customer_tx_summary <- transaction_clean %>%
  group_by(customer_id) %>%
  summarise(
    total_transactions = n(),
    active_transaction_days = n_distinct(transaction_day),
    total_transaction_amount = sum(amount, na.rm = TRUE),
    avg_transaction_amount = mean(amount, na.rm = TRUE),
    median_transaction_amount = median(amount, na.rm = TRUE),
    max_transaction_amount = max(amount, na.rm = TRUE),
    unique_transaction_types = n_distinct(transaction_type),
    .groups = "drop"
  )
```

## Create final Analysis Dataset
```{r}
analysis_data <- customer_clean %>%
  left_join(customer_tx_summary, by = "customer_id")

glimpse(analysis_data)

analysis_data %>%
  summarise(
    total_rows = n(),
    unique_customers = n_distinct(customer_id)
  )
```

# EDA
```{r}
# Churn probability distribution
ggplot(analysis_data, aes(x = churn_probability)) +
  geom_histogram(bins = 30) +
  labs(
    title = "Distribution of Churn Probability",
    x = "Churn Probability",
    y = "Count"
  ) +
  theme_minimal()
```
```{r}
# Customer Lifetime Value
ggplot(analysis_data, aes(x = customer_lifetime_value)) +
  geom_histogram(bins = 30) +
  labs(
    title = "Distribution of Customer Lifetime Value",
    x = "Customer Lifetime Value",
    y = "Count"
  ) +
  theme_minimal()
```
```{r}
# Satisfaction vs churn group
ggplot(analysis_data, aes(x = churn_group, y = satisfaction_score)) +
  geom_boxplot() +
  labs(
    title = "Satisfaction Score by Churn Group",
    x = "Churn Group",
    y = "Satisfaction Score"
  ) +
  theme_minimal()
```
```{r}
# Correlation heatmap for numeric variables
numeric_df <- analysis_data %>%
  select(where(is.numeric)) %>%
  select(
    age, household_size, active_products, app_logins_frequency,
    feature_usage_diversity, tx_count, avg_tx_value, total_tx_volume,
    satisfaction_score, nps_score, support_tickets_count,
    churn_probability, customer_lifetime_value
  )

cor_matrix <- cor(numeric_df, use = "pairwise.complete.obs")

corrplot(cor_matrix, method = "color", type = "upper", tl.cex = 0.8)
```
```{r}
# Transaction trend over time
transaction_clean %>%
  count(transaction_month) %>%
  ggplot(aes(x = transaction_month, y = n)) +
  geom_line() +
  geom_point() +
  labs(
    title = "Monthly Transaction Volume",
    x = "Month",
    y = "Number of Transactions"
  ) +
  theme_minimal()
```

# CDA
```{r}
# Compare customer by churn groups in App
ggplot(analysis_data, aes(x = churn_group, y = app_logins_frequency)) +
  geom_boxplot() +
  labs(
    title = "App Login Frequency by Churn Group",
    x = "Churn Group",
    y = "App Logins Frequency"
  ) +
  theme_minimal()
```
```{r}
# Compare customers by CLV groups on active products
ggplot(analysis_data, aes(x = clv_group, y = active_products)) +
  geom_boxplot() +
  labs(
    title = "Active Products by CLV Group",
    x = "CLV Group",
    y = "Number of Active Products"
  ) +
  theme_minimal()
```
```{r}
# Compare churn group by customer segment
ggplot(analysis_data, aes(x = customer_segment, fill = churn_group)) +
  geom_bar(position = "fill") +
  scale_y_continuous(labels = percent) +
  labs(
    title = "Churn Group Composition by Customer Segment",
    x = "Customer Segment",
    y = "Proportion",
    fill = "Churn Group"
  ) +
  theme_minimal()
```
```{r}
# Compare satisfaction by complaint flag
ggplot(analysis_data, aes(x = complaint_flag, y = satisfaction_score)) +
  geom_boxplot() +
  labs(
    title = "Satisfaction Score by Complaint Status",
    x = "Complaint Flag",
    y = "Satisfaction Score"
  ) +
  theme_minimal()
```
```{r}
# Compare transaction metrics across churn groups
ggplot(analysis_data, aes(x = churn_group, y = total_transactions)) +
  geom_boxplot() +
  labs(
    title = "Total Transactions by Churn Group",
    x = "Churn Group",
    y = "Total Transactions"
  ) +
  theme_minimal()
```
```{r}
analysis_data %>%
  group_by(churn_group) %>%
  summarise(
    avg_app_logins = mean(app_logins_frequency, na.rm = TRUE),
    avg_active_products = mean(active_products, na.rm = TRUE),
    avg_satisfaction = mean(satisfaction_score, na.rm = TRUE),
    avg_nps = mean(nps_score, na.rm = TRUE),
    avg_clv = mean(customer_lifetime_value, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  datatable()
```